---
title: "1. Installation"
description: >
  Alternatives to install the rampage R package. 
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{1. Installation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Stable version from the CRAN


The stable version of the package is available on [CRAN servers](https://cran.r-project.org/package=rampage), which means that you can install it with the regular `install.packages()` function.

```{r cran, eval=FALSE}
install.packages("rampage")
```

If for some reason, the installlation from CRAN fails, you might need to use an alternative method, such as `devtools::install_github("adamkocsis/rampage@main")`


## Stable version from Zenodo (source)

For the sake of completeness and simplicity, the sources of the stable version are also available on GitHub. The `install.packages()` function can be used to install this from source - but you will have to install dependencies of the package.

```{r stable, eval=FALSE}
install.packages(
	"https://zenodo.org/records/15847239/files/rampage_0.2.0.tar.gz?download=1", 
	repos=NULL, type="source")
```

## Development version 

If you encounter technical problems with the package that are not described anywhere, you might want to take a look at the [development version](https://github.com/adamkocsis/rampage/tree/devel).

If you wish to install the development version, I recommend a manual installation:  
1. Clone the repository to your local hard drive.  
2. Open a terminal and navigate to the directory where you cloned. The `rampage` directory should be visible from there.  
3. Run this line in the terminal  

```
R CMD INSTALL rampage
```

<div class="alert alert-danger" role="warning">
- *If you see an error suggesting that `R` is not found, you have to add it to your `PATH` environmental variable.*  
- *If the R packages that `rampage` depend on are not installed, you have to install them manually, or you will get an error.*
</div>


If for some reason, the installlation from CRAN fails, you might need to use an alternative method, such as `devtools::install_github("adamkocsis/rampage@devel")`
