---
title: "Base plots"
description: >
  Ramps with base distribution plotting.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Base plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The purpose of this minimal package is to give full access to the customization and biasing of color ramps in the R environment. The resulting objects can be instantly passed to various plotting methods, such those provided by `sf`, `terra` and `fields` to mention a few.

```{r}
library(rampage)
```

The package itself relies on `grDevices::colorRampPalette` function, which is used to linearly interpolate between colors that are tied to specific values. 

The basic functionality of the package can be combined with any color, either custom-selected or coming from a color ramp function. To provide an example, `rampage` includes a simple 5-color ramp called `gradinv`:

```{r}
gradinv
```

# Histograms

For the sake of a minimal demonstration, `rampage` can be used to colorize bars of a histogram depending on their domain (breakpoints). We can illustrate this with sample from a Gaussian distribution: 

```{r}
set.seed(1)
vals<- rnorm(50000,1,1)
```

## Color tiepoint data frame

The definition of color ramps depend on the given colors (`"color"`) and the the values they are tied to. In the simplest case, we can use 

```{r}
df <- data.frame(
  z=c(-4, 1,  6),
  color=gradinv(3)
)
df
```

Using the default arguments.

```{r}
ex <- expand(df, n=50)
```


This `hist` function can use a set of breakpoints (`breaks`) that will be used to define the bars. The color of the individual bars is vectorized, if you provide exactly as many colors as many bars there are (number of breaks minus one), then the bars will be colored accordingly.

```{r}
hist(vals, breaks=ex$breaks, col=ex$col)
```

## More breaks

The number of levels (breakpoints -1) is directly controled by `n`. A finer resolution can be achieved with 

```{r}
exMore <- expand(df, n=100)
```

```{r}
hist(vals, breaks=exMore$breaks, col=exMore$col)
```


## More tiepoints 

Specifying more tiepoints in the `data.frame` will give more control over the exact placement of the colors. For example if we want to clearly separate the values around 2 (population mean), then we can tie more colors to values that just above and below the values

```{r}
dfMoreTies <- data.frame(
  z=c(-4, 0.5, 1,  1.5, 6),
  color=gradinv(5)
)
dfMoreTies
```

This can then be expanded to a full color ramp:

```{r}
exMoreTies <- rampage::expand(dfMoreTies, n=200)
```

Which can be visualized with the `hist` function.

```{r}
hist(vals, breaks=exMoreTies$breaks, col=exMoreTies$col)
```


The exact placement of the colors are easily customizable by tweaking the `z` values in the `data.frame` that is expanded into the ramp. For instance, if we want the colorings to reflect the values (e.g. 0 is more important than the mean):

```{r}
dfSkewed <- data.frame(
  z=c(-4, -1, 0,  1, 6),
  color=gradinv(5)
)
dfSkewed
```

Expanding to a full color ramp:

```{r}
exSkewed <- rampage::expand(dfSkewed, n=200)
```

```{r}
hist(vals, breaks=exSkewed$breaks, col=exSkewed$col)
```

See the package website for package-specific examples.


