---
title: "2. Topographic maps"
description: >
  Making topographic maps.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{2. Topographic maps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## 1. Topography matrices

The built-in `volcano` topographic data.

```{r}
data(volcano)
```

Attach the package.


### Create a color ramp

```{r}
library(rampage)
```

Specify tiepoints for the color ramp calibration.

```{r}
tiepoints <- data.frame(
	z = c(0, 100, 120, 160,220), 
	color = c("#46AF64", "#96C869", "#E1D791", "#CDB991", "#7D695A")
)
```

Construct an actual color ramp from the tiepoints:

```{r}
# expand
ramp <- expand(tiepoints, n=256) 
str(ramp)
```

Visualize the color ramp.

```{r, plot=TRUE}
plot(ramp)
```

### Plotting with `image` 

Use the color ramp to plot the data.

```{r, fig.height=7, fig.width=7}
image(volcano, breaks=ramp$breaks, col=ramp$col, axes=FALSE)
```

Use the `ramplegend` function to draw a legend:

```{r, fig.height=7, fig.width=7}
image(volcano, breaks=ramp$breaks, col=ramp$col, axes=FALSE)
ramplegend("topright", ramp=ramp, cex=0.7, box=list(border=NA, col="#ffffff55"))
```

### Different shading

```{r}
tiepoints2 <- data.frame(
	z = c(0, 100, 110, 130,200), 
	color = c("#46AF64", "#96C869", "#E1D791", "#CDB991", "#7D695A")
)
ramp2 <- expand(tiepoints2, n=512) 
```

The same topography, same palette, with different tiepoints:

```{r, fig.height=7, fig.width=7}
image(volcano, breaks=ramp2$breaks, col=ramp2$col, axes=FALSE)
ramplegend("topright", ramp=ramp2, cex=0.7, box=list(border=NA, col="#ffffff55"))
```

### Changes

A hypothetical succession assuming a 5 meter increase:

```{r, fig.height=12, fig.width=8}
par(mfrow=c(3,2), mar=c(1,1,1,1))
image(volcano-10, breaks=ramp$breaks, col=ramp$col, axes=FALSE, main="-10 m")
image(volcano-5, breaks=ramp$breaks, col=ramp$col, axes=FALSE, main="-5 m")
image(volcano, breaks=ramp$breaks, col=ramp$col, axes=FALSE, main="+0 m")
image(volcano+5, breaks=ramp$breaks, col=ramp$col, axes=FALSE, main="+5 m")
image(volcano+10, breaks=ramp$breaks, col=ramp$col, axes=FALSE, main="+10 m")
image(volcano+15, breaks=ramp$breaks, col=ramp$col, axes=FALSE, main="+15 m")
ramplegend("bottomright", ramp=ramp, cex=0.7, box=list(border=NA, col="#ffffff55"))
```



### With `imagePlot` 

A similar effect can be reached using the `imagePlot` function from  `fields` extension package, which will display the legend for color ramp automatically.

```{r, fig.height=7, fig.width=7}
library(fields)

# the function call
imagePlot(volcano, breaks=ramp$breaks, col=ramp$col, axes=FALSE)
```




## 2. Georeferenced raster topography

### Built-in objects for topographies


The rampage package includes a built-in set of topogoraphic color schemes (`topos`). These can be accessed with the `data` function.

```{r, echo=TRUE}
data(topos)
```

This object is a `list` of `data.frames` that are potential inputs to the `expand` function. For instance, if you want to generate 256 colors from the `'zagreb'` theme, you can do that with

```{r}
topoRamp <- expand(topos$zagreb, n=256)
```

This ramp can be visualized with the `rampplot` function:

```{r, fig.height=7, fig.width=7}
rampplot(topoRamp)
```

This shows the exact values that are rendered to the specific interals of height.


### Topographies with `terra`

The `terra` extension is an ideal tool to visualize georeference topographic data, such as the [ETOPO](https://www.ncei.noaa.gov/products/etopo-global-relief-model) global topographic relief model. To provide an example, a downscaled 1x1 degree-resolution version of this ETOPO1 version of this model is deposited on the [package's website](http://adamtkocsis.com/rampage/ETOPO1_ice_c_20110606_tiff_1.tif), which you can read in with the following line of code:

```{r, echo=TRUE}
library(terra)
etopo <- rast("http://adamtkocsis.com/rampage/ETOPO1_ice_c_20110606_tiff_1.tif")
etopo
```

This can be visualized with the default viridis palette using the plot function:

```{r, fig.height=7, fig.width=7}
plot(etopo)
```

We can get this to work with the our topographic ramp, by referring to the `col` abd `breaks` arguments of the plot, as we did before.

```{r, fig.height=7, fig.width=7}
plot(etopo, col=topoRamp$col, breaks=topoRamp$breaks)
```

Note that by default this will use a categorical legend, which is not ideal.


```{r, fig.height=7, fig.width=7}
plot(etopo, col=topoRamp$col, breaks=topoRamp$breaks, legend=FALSE)
ramplegend(col=topoRamp$col,breaks=topoRamp$breaks, cex=0.6, x=-160, y=0)
```

Which leaves a much better visual impression.


